%!TEX root = ../main.tex
%----------------------------------------------------------------------
%  Section 5: Data and Baseline Models
%  Paper:   Theory-Informed Kernel Ridge Regression for Asset Pricing
%  Authors: Amedeo Andriollo and Juan F. Imbet
%----------------------------------------------------------------------

\section{Data and Baseline Models}\label{sec:empirical}

%======================================================================
\subsection{Data}\label{sub:data}
%======================================================================

\paragraph{Returns.}
We obtain monthly stock returns from the Center for Research in Security Prices (CRSP) monthly stock file covering 1925--2024, encompassing 38,843 unique securities and 5.16~million security-month observations.
The sample includes ordinary common shares (share codes 10 and 11) listed on NYSE, AMEX, and NASDAQ (exchange codes 1, 2, and 3).
We exclude financial firms (SIC codes 6000--6999) and utilities (SIC codes 4900--4999), whose regulated balance sheets make standard asset-pricing models poor descriptions of their return-generating processes.
We further exclude stocks with share prices below \$5 to mitigate microstructure noise from bid-ask bounce and illiquidity.
Excess returns are computed by subtracting the one-month Treasury bill rate from the Fama-French data library.
The effective sample period begins in July~1963---when both CRSP and Compustat coverage become reliable---and ends in December~2023, spanning 726~months.
This start date matches \citet{gu2020empirical} and the bulk of the cross-sectional return prediction literature.

\paragraph{Firm characteristics.}
We use 209 firm-level predictive characteristics from the Open Source Asset Pricing (OSAP) dataset of \citet{jensen2023machine}, supplemented with three CRSP-derived signals (price, size, and short-term reversal) computed directly from the CRSP monthly file.
The OSAP dataset provides a standardized, publicly available panel of signed characteristics at the security-month level, constructed following the original papers' methodologies and validated against published results.
All characteristics are signed so that higher values predict higher expected returns.
This characteristic set subsumes and extends the 94~characteristics used in \citet{gu2020empirical} and includes size, book-to-market, momentum (2--12 months), short-term reversal, long-term reversal, investment-to-assets, gross profitability, accruals, CAPM beta, idiosyncratic volatility, Amihud illiquidity, earnings surprises, share turnover, and over 190~additional predictors spanning accounting ratios, trading activity, and analyst expectations.

We merge the OSAP characteristics with CRSP returns using the permanent security identifier (permno).
To ensure comparability across variables measured in different units, we cross-sectionally rank each characteristic to the unit interval $[0,1]$ each month.
Ranking is a monotone transformation that preserves the ordinal information in each characteristic while eliminating outliers and ensuring that all variables enter the kernel on a common scale.

The structural restrictions in Section~\ref{sec:restrictions} additionally require firm-level accounting data from Compustat.
We obtain annual fundamentals (525,028 firm-years, 40,678 unique firms, 1950--2025) and quarterly fundamentals (1.87~million firm-quarters, 39,568 unique firms, 1961--2025) from WRDS.
These are merged with CRSP via the CRSP-Compustat linking table using primary link types (LC and LU) with an effective link date filter.
Key accounting variables include total assets, book equity, sales, income before extraordinary items, capital expenditures, property-plant-equipment, depreciation, long-term and short-term debt, cash holdings, R\&D expenditure, and selling-general-administrative expenses.
Our Compustat sample includes both active and inactive firms to avoid survivorship bias.

\paragraph{Macro state variables.}
We augment firm-level characteristics with macroeconomic state variables that serve two roles: they enter the kernel as predictors of expected returns, and they supply inputs to the structural restriction penalties in Section~\ref{sec:restrictions}.
These variables are drawn from four sources.

From the Federal Reserve Economic Data (FRED), we obtain the 3-month Treasury bill rate (TB3MS), the 10-year Treasury yield (GS10), Moody's AAA and BAA corporate bond yields, the Consumer Price Index (CPIAUCSL), industrial production (INDPRO), and the unemployment rate (UNRATE).
The term spread (GS10 minus TB3MS) and the default spread (BAA minus AAA) are computed as differences.
From the National Income and Product Accounts via FRED, we obtain personal consumption expenditures on nondurable goods and services, the PCE price index, and population, which we combine to construct real per capita consumption growth---the central input to the thirteen consumption-based Euler equation restrictions.

The intermediary capital ratio and intermediary capital risk factor of \citet{hekellymanela2017} are obtained from the authors' website at quarterly frequency and interpolated to monthly.
The consumption-wealth ratio $\mathit{cay}$ of \citet{lettauludvigson2001} is constructed following their Dynamic OLS methodology from FRED series on household net worth (TNWBSHNO), employee compensation (COE), and consumption, using 8~leads and lags of the differenced regressors.
The \citet{bakerwurgler2006} investor sentiment index (orthogonalized version) is obtained from the authors' data repository and covers 1965--2023.
The CBOE Volatility Index (VIX) is obtained from FRED and is available from January~1990; we aggregate daily values to monthly averages.
The excess bond premium of \citet{gilchristzakrajsek2012} is proxied by the Moody's BAA spread over the 10-year Treasury (BAAFFM series from FRED).
Additional macro predictors---the dividend-price ratio, earnings-price ratio, aggregate book-to-market, net equity expansion, Treasury bill rate, long-term yield, and stock variance---are obtained from the \citet{welchgoyal2008} dataset, which covers 1871--2024 at monthly frequency.

\paragraph{Missing data and sample restrictions.}
We require a minimum of 15 non-missing characteristics per stock-month observation.
Observations with fewer non-missing characteristics are dropped.
Returns are winsorized at the 0.1st and 99.9th percentiles each month to limit the influence of extreme observations while preserving the bulk of the return distribution.
After applying all filters, the sample contains approximately 2.1~million stock-month observations covering an average of 2,800~stocks per month.

Not all macro state variables are available for the full sample period.
The VIX is available only from 1990, so the two variance and volatility restrictions (44--45) are evaluated on the post-1990 subsample.
The HKM intermediary factors are available quarterly from 1970 to 2012; restrictions~24--27 are evaluated on this subsample.
The Baker-Wurgler sentiment index is available from 1965 to 2023.
All other macro variables cover the full 1963--2023 sample period.

\textbf{[Table 1 about here]}

Table~1 reports summary statistics.
Panel~A presents the distribution of monthly excess returns and selected firm-level characteristics, reporting the mean, standard deviation, and selected percentiles (5th, 25th, 50th, 75th, 95th) across all stock-month observations.
Panel~B reports summary statistics for the macroeconomic state variables.
Panel~C documents sample coverage over time, showing the number of stocks per month across the sample period.

%======================================================================
\subsection{Baseline Models}\label{sub:baselines}
%======================================================================

We compare Theory-KRR against eight baseline models spanning the spectrum from simple linear methods to flexible nonparametric estimators.
All baselines use the same data, the same rolling-window procedure described in Section~\ref{sub:rolling}, and the same LOYO temporal cross-validation for hyperparameter selection.
This ensures that performance differences reflect the models' functional form and regularization, not differences in data or evaluation protocol.

\begin{enumerate}[label=\arabic*.]
\item \textbf{Historical mean.}
The simplest benchmark: $\hat{f}(\mathbf{X}_{i,t}) = \bar{R}$, the pooled cross-sectional average of excess returns in the training sample.
This is the natural benchmark for $R^2_{\mathrm{OOS}}$, which measures improvement over the historical mean.

\item \textbf{OLS.}
Linear regression of next-month excess returns on the full vector of characteristics: $\hat{f}(\mathbf{X}_{i,t}) = \hat{\bm{\beta}}^{\top}\mathbf{X}_{i,t}$.
With 37 characteristics, OLS is prone to overfitting in small training samples but provides a benchmark for the value of nonlinearity.

\item \textbf{Ridge regression.}
OLS with an $\ell_2$ penalty on the coefficient vector: the objective is $\frac{1}{n}\sum_s(R_s - \bm{\beta}^\top \mathbf{X}_s)^2 + \lambda\|\bm{\beta}\|_2^2$.
Ridge regression shrinks coefficients toward zero without enforcing sparsity.
The regularization parameter $\lambda$ is selected by LOYO CV.

\item \textbf{Lasso.}
OLS with an $\ell_1$ penalty: the objective is $\frac{1}{n}\sum_s(R_s - \bm{\beta}^\top \mathbf{X}_s)^2 + \lambda\|\bm{\beta}\|_1$.
The $\ell_1$ penalty performs variable selection by setting some coefficients exactly to zero.

\item \textbf{Elastic net.}
A convex combination of $\ell_1$ and $\ell_2$ penalties: the objective is $\frac{1}{n}\sum_s(R_s - \bm{\beta}^\top \mathbf{X}_s)^2 + \lambda_1\|\bm{\beta}\|_1 + \lambda_2\|\bm{\beta}\|_2^2$.
The elastic net combines variable selection (Lasso) with coefficient shrinkage (ridge).
Both $\lambda_1$ and $\lambda_2$ are selected by LOYO CV.

\item \textbf{Standard KRR.}
The Gaussian kernel ridge regression estimator with RKHS-norm regularization only: this is Theory-KRR with all $\mu_j = 0$.
Standard KRR captures nonlinearities and interactions through the Gaussian kernel but receives no guidance from economic theory.
Comparing Theory-KRR to standard KRR isolates the marginal contribution of the structural penalties.

\item \textbf{Neural network.}
A feedforward neural network with three hidden layers, following the architecture of Gu, Kelly, and Xiu~(2020): layer widths of 32, 16, and 8 neurons, ReLU activation functions, batch normalization, and dropout ($p = 0.5$) for regularization.
The network is trained by stochastic gradient descent with early stopping based on validation loss (the last 20\% of the training window serves as the validation set).

\item \textbf{Random forest.}
An ensemble of 300 regression trees with the number of candidate variables at each split set to $\lfloor p/3 \rfloor$, following Gu, Kelly, and Xiu~(2020).
Tree depth is controlled by requiring a minimum of 5 observations per leaf node.
\end{enumerate}

%======================================================================
\subsection{Evaluation Metrics}\label{sub:evaluation}
%======================================================================

We evaluate predictive performance along three dimensions: statistical accuracy, portfolio profitability, and economic value.

\paragraph{Out-of-sample $R^2$.}
Following Campbell and Thompson~(2008), we measure predictive accuracy by
\begin{equation}\label{eq:r2oos}
R^2_{\mathrm{OOS}} = 1 - \frac{\displaystyle\sum_{(i,t) \in \mathcal{T}_{\mathrm{OOS}}} \bigl(R_{i,t+1} - \hat{f}(\mathbf{X}_{i,t})\bigr)^2}{\displaystyle\sum_{(i,t) \in \mathcal{T}_{\mathrm{OOS}}} \bigl(R_{i,t+1} - \bar{R}\bigr)^2},
\end{equation}
where $\mathcal{T}_{\mathrm{OOS}}$ indexes the out-of-sample observations and $\bar{R}$ is the historical mean computed from the training sample.
A positive $R^2_{\mathrm{OOS}}$ indicates that the model predicts better than the historical mean.
In the cross-section of individual stock returns, $R^2_{\mathrm{OOS}}$ values above 1\% are economically meaningful given the high idiosyncratic variance of monthly returns.

\paragraph{Portfolio Sharpe ratios.}
We form decile portfolios each month by sorting stocks on their predicted expected returns $\hat{f}(\mathbf{X}_{i,t})$.
The long-short portfolio goes long the top decile and short the bottom decile, with value-weighted holdings within each decile.
We report annualized Sharpe ratios:
\begin{equation}\label{eq:sharpe}
\mathrm{SR} = \frac{\hat{\mu}_{L-S}}{\hat{\sigma}_{L-S}} \times \sqrt{12},
\end{equation}
where $\hat{\mu}_{L-S}$ and $\hat{\sigma}_{L-S}$ are the sample mean and standard deviation of the monthly long-short return.
Higher Sharpe ratios indicate that the model's predictions generate more profitable trading strategies per unit of risk.

\paragraph{Certainty equivalent return.}
For a mean-variance investor with relative risk aversion $\gamma = 5$, the certainty equivalent return of a portfolio with mean $\hat{\mu}_p$ and variance $\hat{\sigma}_p^2$ is
\begin{equation}\label{eq:cer}
\mathrm{CER} = \hat{\mu}_p - \frac{\gamma}{2}\,\hat{\sigma}_p^2.
\end{equation}
The CER provides an economically interpretable measure of the value of predictability: it is the riskless return that the investor would accept in lieu of the risky portfolio.
We compute the CER for the long-short portfolio constructed from each model's predictions.

\paragraph{Statistical inference.}
We assess the statistical significance of performance differences using three procedures.
The Diebold-Mariano test compares squared forecast errors between Theory-KRR and each baseline, with Newey-West standard errors (6 lags) to account for serial correlation.
For Sharpe ratio differences, we use a block bootstrap with 12-month blocks and 10,000 replications; the block length preserves the annual dependence structure in monthly returns.
All reported standard errors use the Newey-West HAC estimator with 6 lags unless otherwise noted.

%======================================================================
\subsection{Summary Statistics}\label{sub:summary}
%======================================================================

\textbf{[Table 1 about here]}

Table~1 documents the key features of our dataset.
Panel~A shows that the average monthly excess return is 0.9\% with a standard deviation of 14.2\%, consistent with well-known properties of the cross-section.
Firm characteristics exhibit substantial cross-sectional dispersion: for example, the interquartile range of log market equity spans roughly two orders of magnitude, and the book-to-market ratio ranges from 0.12 at the 5th percentile to 2.31 at the 95th percentile.
Panel~B reports that macroeconomic state variables display moderate persistence and time variation; the intermediary capital ratio of He, Kelly, and Manela~(2017) has an autocorrelation coefficient of 0.97, while the Baker-Wurgler sentiment index has an autocorrelation of 0.92.
Panel~C confirms that sample coverage increases over time, from approximately 1,500 stocks per month in the 1960s to over 3,500 stocks per month in the 2000s, with a decline to approximately 3,000 after 2010 as the number of publicly listed U.S.\ firms decreases.
