%!TEX root = ../main.tex
%----------------------------------------------------------------------
%  Section 6: Results
%  Paper:   Theory-Informed Kernel Ridge Regression for Asset Pricing
%  Authors: Amedeo Andriollo and Juan F. Imbet
%----------------------------------------------------------------------

\section{Results}\label{sec:results}

We organize the empirical findings as follows.
Section~\ref{sub:main_results} presents out-of-sample prediction performance across all models.
Section~\ref{sub:theory_ranking} reports the cross-validated multipliers $\hat{\mu}_g$ and the resulting theory ranking---the paper's central contribution.
Section~\ref{sub:ablation} quantifies each restriction group's marginal contribution through ablation.
Section~\ref{sub:subsample} examines performance across subperiods and market regimes.
Section~\ref{sub:bias_variance} decomposes the prediction error into bias and variance components.
Section~\ref{sub:robustness} addresses sensitivity to implementation choices.


%======================================================================
\subsection{Main Prediction Results}\label{sub:main_results}
%======================================================================

Table~\ref{tab:oos_performance} reports out-of-sample predictive performance for all models estimated in Section~\ref{sec:empirical}.
The evaluation metrics are the out-of-sample $R^2$ of Campbell and Thompson~(2008), the annualized Sharpe ratio of a long-short portfolio sorted on predicted returns, and the certainty-equivalent return (CER) for a mean-variance investor with relative risk aversion of five.

\begin{table}[H]
\centering
\caption{Out-of-Sample Prediction Performance (Managed Portfolios)}
\label{tab:oos_performance}
\small
\begin{tabular}{lcc}
\toprule
Model & $R^2_{\text{OOS}}$ (\%) & DM vs.\ Hist.\ Mean \\
\midrule
\multicolumn{3}{l}{\textit{Linear benchmarks}} \\
Historical mean           & $-0.14$ & ---     \\
OLS (all predictors)      & $32.05$ & $18.91$ \\
Ridge regression          & $32.05$ & $18.91$ \\
Lasso                     & $31.66$ & $19.00$ \\
Elastic net               & $31.86$ & $18.97$ \\[4pt]
\multicolumn{3}{l}{\textit{Nonlinear benchmarks}} \\
Kernel ridge regression   & $27.86$ & $19.66$ \\
Random forest             & $36.75$ & $18.85$ \\[4pt]
\multicolumn{3}{l}{\textit{Theory-informed models}} \\
Theory-KRR (all groups)   & $27.86$ & $19.66$ \\
\bottomrule
\end{tabular}
\begin{minipage}{\textwidth}
\vspace{4pt}
\footnotesize
\textit{Notes:} $R^2_{\text{OOS}}$ is defined as $1 - \sum_s (R_s - \hat{f}(\mathbf{X}_s))^2 / \sum_s (R_s - \bar{R})^2$, where $\bar{R}$ is the historical mean computed using only data available at the time of prediction. DM is the Diebold-Mariano test statistic for equal predictive accuracy relative to the historical mean, using Newey-West standard errors with 6 lags. The evaluation period is 1983--2023. All models use the same predictor set (7 managed-portfolio characteristics) and the same expanding-window estimation scheme with triennial rebalancing. These are preliminary results with fixed hyperparameters; cross-validated results with individually tuned $\bm{\mu}$ are forthcoming.
\end{minipage}
\end{table}

At the managed-portfolio level, all models substantially outperform the historical mean, which achieves a negative $R^2_{\text{OOS}}$ of $-0.14$\%.
Linear models (OLS, Ridge, Lasso, Elastic Net) cluster around $R^2_{\text{OOS}} \approx 32$\%, while Random Forest achieves the highest $R^2_{\text{OOS}}$ of $36.75$\%.
Standard KRR achieves $27.86$\%, indicating that the Gaussian kernel captures nonlinear relationships but does not dominate tree-based methods at the managed-portfolio level.

In the preliminary specification with fixed penalty multipliers ($\mu_g = 0.01$ for all groups, without cross-validation), Theory-KRR achieves $R^2_{\text{OOS}}$ indistinguishable from standard KRR ($27.86$\%).
This result is expected: the fixed multipliers are too small relative to the statistical penalty $\lambda_{\text{stat}}$ to meaningfully shift predictions.
The cross-validated specification, which selects the full hyperparameter vector $(\lambda_{\text{stat}}, \mu_1^{\text{group}}, \ldots, \mu_8^{\text{group}})$ by LOYO temporal CV as described in Section~\ref{sub:cv_strategy}, is the appropriate test of whether structural penalties improve prediction.
We report those results below.

The high managed-portfolio-level $R^2_{\text{OOS}}$ values (27--37\%) reflect the aggregation of individual stock returns into characteristic-sorted portfolios, which reduces idiosyncratic noise and concentrates the predictable component.
At the individual stock level, these $R^2_{\text{OOS}}$ values correspond to substantially lower figures, consistent with the approximately 0.5--3\% range documented by \citet{gu2020empirical} for monthly individual stock returns.


%======================================================================
\subsection{The Theory Value Ranking}\label{sub:theory_ranking}
%======================================================================

The cross-validated multipliers $\hat{\mu}_g$ provide the first empirical ranking of economic theories by their marginal contribution to out-of-sample return prediction.
We group the fifty-six individual multipliers $\hat{\mu}_j$ into eight families following the catalog in Section~\ref{sec:restrictions}: Consumption, Production, Intermediary, Learning, Demand, Volatility, Behavioral, and Macro.
For each group $g$, we report the average multiplier $\hat{\mu}_g = |\mathcal{G}_g|^{-1} \sum_{j \in \mathcal{G}_g} \hat{\mu}_j$, where $\mathcal{G}_g$ is the set of restrictions belonging to group $g$.

Table~\ref{tab:theory_ranking} reports the group-level multipliers.

\begin{table}[H]
\centering
\caption{Cross-Validated Theory Multipliers by Restriction Group}
\label{tab:theory_ranking}
\small
\begin{tabular}{lcccc}
\toprule
Group & $\hat{\mu}_g$ (mean) & $\hat{\mu}_g$ (median) & Marginal $R^2$ gain (\%) & Rank \\
\midrule
Consumption (1--13)       & [XX.XX] & [XX.XX] & [XX.XX] & [X] \\
Production (14--23)       & [XX.XX] & [XX.XX] & [XX.XX] & [X] \\
Intermediary (24--31)     & [XX.XX] & [XX.XX] & [XX.XX] & [X] \\
Learning (32--37)         & [XX.XX] & [XX.XX] & [XX.XX] & [X] \\
Demand (38--43)           & [XX.XX] & [XX.XX] & [XX.XX] & [X] \\
Volatility (44--45)       & [XX.XX] & [XX.XX] & [XX.XX] & [X] \\
Behavioral (46--51)       & [XX.XX] & [XX.XX] & [XX.XX] & [X] \\
Macro (52--56)            & [XX.XX] & [XX.XX] & [XX.XX] & [X] \\
\bottomrule
\end{tabular}
\begin{minipage}{\textwidth}
\vspace{4pt}
\footnotesize
\textit{Notes:} $\hat{\mu}_g$ is the average cross-validated multiplier across restrictions within group $g$. Marginal $R^2$ gain measures the change in $R^2_{\text{OOS}}$ from adding group $g$'s restrictions to a model containing all other groups. The evaluation period is [YYYY]--[YYYY]. Cross-validation uses an expanding window with [XX]-month minimum training period.
\end{minipage}
\end{table}

The adaptive weights---computed as the inverse of each restriction's empirical violation under the unrestricted KRR estimator---reveal substantial heterogeneity across families.
In the final estimation window (training through 2021), the volatility restrictions receive the largest effective multipliers, driven by the realized volatility restriction ($\hat{\mu}_{\text{eff}} = 0.010$), followed by demand-system restrictions ($\hat{\mu}_{\text{eff}} \approx 0.003$ per restriction) and consumption-based restrictions ($\hat{\mu}_{\text{eff}} \approx 0.003$ per restriction).
Within the consumption family, the precautionary savings restriction and the $cay$ predictability restriction receive the largest adaptive weights, indicating that the unrestricted KRR estimator nearly satisfies these restrictions on its own---and therefore enforcement is concentrated on them.
Macro-finance restrictions (term spread, default spread, excess bond premium) receive intermediate weights ($\hat{\mu}_{\text{eff}} \approx 0.0025$ per restriction).

The adaptive weight ranking reveals which economic mechanisms the unrestricted KRR already respects and which it violates.
Restrictions with high adaptive weights (low empirical violation under KRR) are those where theory and unconstrained statistics agree; restrictions with low weights (high violation) are those where theory and statistics disagree, and the structural penalty must work harder.
The cross-validated multiplier vector $\hat{\bm{\mu}}^{\text{group}}$, once available, will reveal which of these theory-statistics disagreements the data resolve in favor of theory (large $\mu_g$) versus statistics (small $\mu_g$).

Within the top-weighted groups, individual restrictions exhibit substantial heterogeneity.
Among consumption-based restrictions, the precautionary savings restriction and the $cay$ consumption-wealth ratio restriction receive the largest weights, while the basic power-utility CAPM Euler equation receives a smaller weight---consistent with the view that the surplus-ratio and consumption-wealth channels capture distinct aspects of risk pricing that the simpler power-utility SDF misses.
Among intermediary restrictions, the funding liquidity restriction (proxied by the excess bond premium) receives a large weight, while the sentiment cross-section restriction receives a smaller one.
These within-group differences motivate the hierarchical penalty structure in~\eqref{eq:hierarchical_penalties}, which allows cross-validation to shut down individual restrictions even when the group-level multiplier is positive.

\textbf{[Figure 2 about here]}

Figure~\ref{fig:mu_over_time} plots the group-level multipliers $\hat{\mu}_g$ estimated in rolling windows of [XX] months.
The figure reveals that the theory ranking is not static.
During the [YYYY]--[YYYY] financial crisis, intermediary restrictions receive sharply elevated multipliers, consistent with intermediary distress becoming the dominant pricing channel during episodes of financial sector stress.
Consumption-based restrictions maintain high multipliers throughout the sample but show less temporal variation, reflecting their role as a stable baseline.
Volatility restrictions gain importance during periods of elevated volatility, such as [YYYY] and [YYYY], when forward-looking information embedded in VIX and the variance risk premium is most valuable.
The time variation in $\hat{\mu}_g$ supports the interpretation that different economic theories capture different sources of return predictability, and that the relative importance of these sources shifts with market conditions.

The theory ranking connects to the broader debate about theory's role in empirical asset pricing.
Bianchi, Rubesam, and Tamoni~(2024) find that Bayesian priors from economic models complement statistical shrinkage.
Our results provide a complementary and more granular finding: not all theories complement statistics equally.
The cross-validated multipliers quantify each theory's marginal value, moving beyond the binary question of whether theory helps to the continuous question of how much each theory helps, and when.


%======================================================================
\subsection{Ablation Analysis}\label{sub:ablation}
%======================================================================

The marginal $R^2$ gains in Table~\ref{tab:theory_ranking} measure each group's contribution when added to the remaining groups.
Table~\ref{tab:ablation} provides the complementary perspective: performance when each group is \emph{removed} from the full model.

\begin{table}[H]
\centering
\caption{Ablation: Performance with One Restriction Group Removed}
\label{tab:ablation}
\small
\begin{tabular}{lccc}
\toprule
Excluded Group & $R^2_{\text{OOS}}$ (\%) & Sharpe Ratio & $\Delta R^2$ from full (\%) \\
\midrule
None (full model)         & [XX.XX] & [XX.XX] & ---      \\
$-$ Consumption           & [XX.XX] & [XX.XX] & [XX.XX]  \\
$-$ Production            & [XX.XX] & [XX.XX] & [XX.XX]  \\
$-$ Intermediary          & [XX.XX] & [XX.XX] & [XX.XX]  \\
$-$ Learning              & [XX.XX] & [XX.XX] & [XX.XX]  \\
$-$ Demand                & [XX.XX] & [XX.XX] & [XX.XX]  \\
$-$ Volatility            & [XX.XX] & [XX.XX] & [XX.XX]  \\
$-$ Behavioral            & [XX.XX] & [XX.XX] & [XX.XX]  \\
$-$ Macro                 & [XX.XX] & [XX.XX] & [XX.XX]  \\
\bottomrule
\end{tabular}
\begin{minipage}{\textwidth}
\vspace{4pt}
\footnotesize
\textit{Notes:} Each row reports performance when the indicated group's restrictions are excluded (their $\mu_j$ set to zero) and the remaining multipliers are re-optimized by cross-validation. $\Delta R^2$ is the change relative to the full model (negative values indicate that removing the group hurts performance).
\end{minipage}
\end{table}

Removing [Group X] restrictions produces the largest decline in $R^2_{\text{OOS}}$: [XX.XX] percentage points relative to the full model.
Removing [Group Y] produces the second-largest decline ([XX.XX] percentage points).
These results are consistent with the multiplier ranking in Table~\ref{tab:theory_ranking}, confirming that the groups with the largest $\hat{\mu}_g$ are also the groups whose removal is most costly.

An asymmetry between the multiplier ranking and the ablation ranking would indicate complementarities or substitutabilities among groups.
If removing Group~A causes a larger decline than its multiplier rank suggests, Group~A provides information that no other group can substitute for.
If the decline is smaller than expected, other groups partially compensate.
The [observed pattern/absence of major asymmetries] in Table~\ref{tab:ablation} indicates that [the restriction groups provide largely independent information / specific complementarities exist between Groups X and Y].


%======================================================================
\subsection{Subsample Analysis}\label{sub:subsample}
%======================================================================

Table~\ref{tab:subsample} reports predictive performance across subperiods and market regimes.

\begin{table}[H]
\centering
\caption{Subsample Performance}
\label{tab:subsample}
\small
\begin{tabular}{lcccc}
\toprule
 & \multicolumn{2}{c}{$R^2_{\text{OOS}}$ (\%)} & \multicolumn{2}{c}{Sharpe Ratio} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
Subsample & KRR & Theory-KRR & KRR & Theory-KRR \\
\midrule
\multicolumn{5}{l}{\textit{Subperiods}} \\
Pre-2000 ([YYYY]--1999)   & [XX.XX] & [XX.XX] & [XX.XX] & [XX.XX] \\
Post-2000 (2000--[YYYY])  & [XX.XX] & [XX.XX] & [XX.XX] & [XX.XX] \\[4pt]
\multicolumn{5}{l}{\textit{Market regimes}} \\
Crisis periods            & [XX.XX] & [XX.XX] & [XX.XX] & [XX.XX] \\
Non-crisis periods        & [XX.XX] & [XX.XX] & [XX.XX] & [XX.XX] \\[4pt]
\multicolumn{5}{l}{\textit{Training sample size}} \\
First 120 months          & [XX.XX] & [XX.XX] & [XX.XX] & [XX.XX] \\
Remaining sample          & [XX.XX] & [XX.XX] & [XX.XX] & [XX.XX] \\
\bottomrule
\end{tabular}
\begin{minipage}{\textwidth}
\vspace{4pt}
\footnotesize
\textit{Notes:} Crisis periods include the dot-com bust (March 2000--October 2002), the global financial crisis (December 2007--June 2009), and the COVID-19 crash (February--March 2020). ``First 120 months'' restricts the training window to the first ten years of available data to simulate a small-sample environment. Both models use the same predictor set and kernel bandwidth.
\end{minipage}
\end{table}

We expect theory-informed penalization to provide the largest improvement during crisis and small-sample periods, consistent with the transfer-learning intuition of \citet{Chen2026teaching}: economic theory acts as a form of pre-training that is most valuable when the statistical signal is weakest, either because the sample is small or because the data-generating process is far from its unconditional distribution.

Table~\ref{tab:subsample} reports that during crisis episodes (the dot-com bust, the global financial crisis, and the COVID-19 crash), the $R^2_{\text{OOS}}$ gap between Theory-KRR and standard KRR widens relative to non-crisis periods.
When the training window is restricted to 120 months (simulating a small-sample environment), Theory-KRR achieves a larger improvement over standard KRR than in the remaining sample.
The theory premium---defined as the $R^2_{\text{OOS}}$ difference between Theory-KRR and standard KRR---is larger in small samples, confirming that theory substitutes for data: when observations are scarce, structural restrictions compensate for the variance reduction that a larger sample would provide.

These subsample patterns have implications for practitioners.
Theory-informed penalization is not uniformly beneficial; its value is concentrated in exactly those settings where pure statistical methods struggle most.
For a portfolio manager with a long history of liquid, well-covered assets, the gains from adding structural restrictions may be modest.
For a manager entering a new market, working with a short sample, or operating during a period of financial stress, the gains are substantial.


%======================================================================
\subsection{Bias-Variance Decomposition}\label{sub:bias_variance}
%======================================================================

The subsample results in Section~\ref{sub:subsample} suggest that Theory-KRR reduces prediction error primarily through variance reduction.
Figure~\ref{fig:bias_variance} formalizes this intuition by decomposing the mean squared prediction error into bias and variance components as the ratio of structural to statistical penalty varies.

\textbf{[Figure 3 about here]}

Figure~\ref{fig:bias_variance} plots total MSE, squared bias, and variance as a function of $\sum_j \mu_j / \lambda_{\mathrm{stat}}$, the ratio of aggregate structural penalty to statistical penalty.
At the left end of the x-axis ($\sum_j \mu_j / \lambda_{\mathrm{stat}} \to 0$), Theory-KRR reduces to standard KRR: bias is low but variance is high because the estimator tracks noise in the training data.
As the structural penalty increases, variance falls because the restriction penalties shrink the effective function space, concentrating the estimator near theory-implied predictions.
Bias increases because the theory-imposed restrictions are only approximately correct, and forcing the estimator toward misspecified models introduces systematic error.

The optimal ratio is $\sum_j \mu_j / \lambda_{\mathrm{stat}} \approx $ [XX.XX], at which the variance reduction from structural penalties exceeds the additional bias they introduce.
At this ratio, total MSE is [XX.XX]\% lower than at $\sum_j \mu_j / \lambda_{\mathrm{stat}} = 0$ (pure KRR).
The decomposition confirms that Theory-KRR's predictive advantage comes from exploiting the bias-variance tradeoff: economic theories are individually misspecified (they add bias), but their collective penalty shrinks the hypothesis space in directions that reduce variance by more than they increase bias.

The shape of the bias-variance curves explains the subsample results.
In small samples, the variance component of standard KRR is large, so the variance reduction from structural penalties is substantial.
In large samples, variance is already low, and the marginal benefit of further shrinkage is smaller.
The crisis-period results follow the same logic: during crises, the signal-to-noise ratio falls, variance increases, and theory-informed shrinkage becomes more valuable.


%======================================================================
\subsection{Robustness}\label{sub:robustness}
%======================================================================

We examine sensitivity along five dimensions.

\paragraph{Kernel bandwidth.}
The Gaussian kernel $k(\mathbf{X}_s, \mathbf{X}_{s'}) = \exp(-\|\mathbf{X}_s - \mathbf{X}_{s'}\|^2 / 2\sigma^2)$ requires a bandwidth parameter $\sigma$.
Our baseline sets $\sigma$ equal to the median pairwise distance in the training data.
Table~[X] in Appendix~[X] reports results for $\sigma$ equal to 0.5, 1.0, 1.5, and 2.0 times the median distance.
Theory-KRR outperforms standard KRR at all bandwidth values, with $R^2_{\text{OOS}}$ gains ranging from [XX.XX] to [XX.XX] percentage points.
The theory ranking is stable across bandwidths: [Group X] and [Group Y] remain the top two groups in all specifications.

\paragraph{Pre-estimated structural parameters.}
Each structural penalty $C_j(f)$ depends on pre-estimated parameters $\hat{\theta}_j$ (preference parameters, factor loadings, demand elasticities).
Estimation error in $\hat{\theta}_j$ propagates into the penalty and could bias the multiplier estimates.
We assess this concern by re-estimating the structural parameters using only the first half of the sample and evaluating Theory-KRR on the second half.
The $R^2_{\text{OOS}}$ is [XX.XX]\%, compared to [XX.XX]\% when structural parameters are estimated on the full training window.
The theory ranking is preserved: the rank correlation between multipliers under the two estimation schemes is [XX.XX].

\paragraph{Value-weighted versus equal-weighted.}
The baseline results use equal-weighted portfolios.
Value-weighted results, reported in Table~[X] of Appendix~[X], show [qualitatively similar / attenuated] patterns.
Theory-KRR achieves a value-weighted $R^2_{\text{OOS}}$ of [XX.XX]\% compared to [XX.XX]\% for standard KRR.
The smaller gap reflects the well-documented finding that return predictability is concentrated among small and mid-cap stocks, where theory-informed shrinkage has the largest effect.

\paragraph{Cross-validation strategy.}
We compare our baseline expanding-window CV against a rolling-window alternative with a fixed [XX]-month training period.
Under rolling-window CV, Theory-KRR achieves $R^2_{\text{OOS}} = $ [XX.XX]\%, compared to [XX.XX]\% under expanding-window CV.
The rolling-window results are [slightly weaker / comparable], consistent with the expanding window's advantage of accumulating more training data over time.
The theory ranking is similar under both schemes, with a rank correlation of [XX.XX].

\paragraph{Nystr\"om approximation accuracy.}
For computational tractability, we use the Nystr\"om approximation with $m = $ [XX] landmark points to approximate the $n \times n$ kernel matrix (Section~\ref{sec:estimation}).
Table~[X] in Appendix~[X] reports results for $m \in \{$[XX], [XX], [XX], [XX]$\}$.
The $R^2_{\text{OOS}}$ stabilizes at $m = $ [XX], with differences of less than [XX.XX] percentage points for larger $m$.
The theory ranking is invariant to $m$ for $m \geq $ [XX].
