% Theories as Regularizers
% Amedeo Andriollo and Juan F. Imbet
% Paris Dauphine University - PSL

\documentclass[12pt]{article}

% ── Page layout ──────────────────────────────────────────────
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\doublespacing

% ── Mathematics ──────────────────────────────────────────────
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{bm}

% ── Tables and figures ───────────────────────────────────────
\usepackage{booktabs,tabularx,multirow,array}
\usepackage{graphicx}
\usepackage{caption,subcaption}

% ── Bibliography ─────────────────────────────────────────────
\usepackage[authoryear,round]{natbib}
\bibliographystyle{plainnat}

% ── Cross-references and links ───────────────────────────────
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  citecolor=blue!60!black,
  urlcolor=blue!60!black,
  pdfauthor={Amedeo Andriollo and Juan F. Imbet},
  pdftitle={Theories as Regularizers}
}
\usepackage{cleveref}

% ── Miscellaneous ────────────────────────────────────────────
\usepackage{enumerate,enumitem}
\usepackage{xcolor}
\usepackage{dsfont}           % indicator function \mathds{1}

% ── Theorem environments ────────────────────────────────────
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% ── Custom commands ──────────────────────────────────────────
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Hcal}{\mathcal{H}}
\newcommand{\Kcal}{\mathcal{K}}
\newcommand{\Fcal}{\mathcal{F}}
\newcommand{\ind}{\mathds{1}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

% ══════════════════════════════════════════════════════════════
%  Title
% ══════════════════════════════════════════════════════════════

\title{\textbf{Theories as Regularizers}\thanks{%
We thank [acknowledgments].
Andriollo: Paris Dauphine University -- PSL, \texttt{amedeo.andriollo@dauphine.psl.eu}.
Imbet: Paris Dauphine University -- PSL, \texttt{juan.imbet@dauphine.psl.eu}.}}
\author{Amedeo Andriollo \and Juan F.\ Imbet}
\date{\today}

\begin{document}

\maketitle

% ── Abstract ─────────────────────────────────────────────────
\begin{abstract}
\begin{singlespace}
\noindent
Machine-learning methods for return prediction ignore economic theory. Structural asset-pricing models encode deep knowledge about expected returns but are individually misspecified. We let the data decide which theories help and how much. We embed over fifty model-implied restrictions---from consumption-based, production-based, intermediary, behavioral, and demand-system frameworks---as soft penalties in a kernel ridge regression estimator. Each restriction carries its own penalty weight, selected by cross-validation: theories that hurt prediction are silenced; theories that help are amplified. The result is the first empirical ranking of economic theories by their marginal out-of-sample forecasting value. Theory-informed penalization delivers higher out-of-sample predictive accuracy and Sharpe ratios than both agnostic machine learning and standard kernel methods. Consumption-based and intermediary restrictions contribute the most. Theory helps disproportionately when data are scarce and during crises---precisely when investors need guidance most.

\medskip
\noindent\textbf{JEL Classification:} G12, G17, C14, C52.

\noindent\textbf{Keywords:} Asset pricing, kernel ridge regression, structural restrictions, machine learning, model selection.
\end{singlespace}
\end{abstract}

\thispagestyle{empty}
\newpage
\setcounter{page}{1}

% ══════════════════════════════════════════════════════════════
%  Main body
% ══════════════════════════════════════════════════════════════

\input{sections/introduction}
\input{sections/framework}
\input{sections/restrictions}
\input{sections/estimation}
\input{sections/empirical}
\input{sections/results}
\input{sections/conclusion}

% ══════════════════════════════════════════════════════════════
%  References
% ══════════════════════════════════════════════════════════════

\newpage
\bibliography{references}

% ══════════════════════════════════════════════════════════════
%  Appendix
% ══════════════════════════════════════════════════════════════

\newpage
\input{tables_and_figures}

% ══════════════════════════════════════════════════════════════════════
%  Appendix
% ══════════════════════════════════════════════════════════════════════

\newpage
\appendix
\input{appendix/appendix}

\end{document}
